---
title: "04 - Calculate Overlap"
author: "Stuart Brown"
output: 
  html_document: 
    theme: flatly
    toc: yes
    toc_depth: 3
    toc_float: true
    fig_width: 7
    fig_height: 5
    fig_caption: yes
    number_sections: no
    keep_md: yes
    df_print: paged
    code_folding: show
    highlight: tango
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, rows.print = 10, cols.print = 10,
                      rownames.print = FALSE, warning = FALSE,
                      message = FALSE, eval = TRUE, results = "markup", 
                      error = FALSE, strip.white = TRUE,
                      tidy = "styler", dpi = 320, fig.align = "center",
                      strip.white = TRUE, 
                      dev.args = list(png  = list(type = "cairo")))
```

## Read in the data

Here we read in the climate data which we used to determine overlap. A number of other datasets are read in for plotting etc.

```{r importData, results='markup'}
library(data.table)
library(Hmisc)
library(terra)
library(sf)
library(units)
library(ggplot2)
library(gtools)
library(rnaturalearth)
library(overlapping)
library(cowplot)
library(ggpattern)
library(patchwork)
library(magick)
library(cptcity)

# function to round to arbitrary number
round_any <- function(x, accuracy, f = round) {f(x / accuracy) * accuracy}

source("source/rotate_prj.R")

# Coastlines and aesthetics
coastlines <- ne_coastline(110, returnclass = "sf")
land <- ne_countries(110, returnclass = "sf")
oceans <- read_sf("outputs/output_shapes.gpkg", layer = "ne_ocean")

# Marine realms from Costello et al
regions <- read_sf("outputs/output_shapes.gpkg", layer = "marine_regions_clipped")

st_agr(oceans) <- "constant"
st_agr(regions) <- "constant"

# Area of each region in km^2
regions$Area <- set_units(st_area(regions), "km^2")

# 2.5° Ocean mask
ocean_mask <- rast("outputs/ocean_mask.grd")

# Climate data masked to oceans only
## RCP
r45 <- mask(rast("outputs/rcp45_rasters_ocean_warming.grd")[["stSNR_ocn_warm"]], 
            mask = ocean_mask)
r85 <-  mask(rast("outputs/rcp85_rasters_ocean_warming.grd")[["stSNR_ocn_warm"]], 
            mask = ocean_mask)
rcp_snr <- rast(list(r45, r85))
names(rcp_snr) <- c("RCP45", "RCP85")

## Paleo
paleo_snr <- mask(rast("outputs/paleo_rasters_ocean_warming.grd")[["stSNR_ocn_warm"]], 
            mask = ocean_mask)
names(paleo_snr) <- "TraCE"

## combined
snr_comb <- rast(list(paleo_snr, rcp_snr))
snr_comb
```

A plot of the climate data shows us the following (we saw a similar plot in **01 - StableClim Extract**)

```{r plotSNR, echo = FALSE, fig.show='asis', fig.cap="SNR values for the past and two future scenarios", fig.align='center', fig.height=5, fig.width=9}
xdf <- melt(setDT(as.data.frame(rast(list(paleo_snr, rcp_snr)),
                                xy = TRUE)), id.vars = c("x", "y"))
xdf[, variable := factor(variable, levels = c("TraCE", "RCP45", "RCP85"))]
ggplot() +
  geom_raster(data = xdf, aes(x, y, fill = value)) +
  facet_wrap(~variable, ncol = 2) +
  scale_fill_gradientn(colours = hcl.colors(100, "Temps"),
                       limits = c(0, 6.0),
                       labels = c("Low","", "High"),
                       breaks = c(0, 3, 6.0),
                       oob = scales::squish,
                       guide = guide_colourbar(
                         title = "SNR",
                         title.position = "top",
                         title.hjust = 0.5,
                         title.theme = element_text(size = 12, colour = "black"),
                         label.position = "bottom",
                         label.hjust = 0.5,
                         label.theme = element_text(size = 12, colour = "black"),
                         barwidth = 10,
                         barheight = 0.5,
                         nbin = 100,
                         frame.colour = "black",
                         frame.linewidth = 0.5
                       )) +
  geom_sf(data = land, colour = NA, fill = "grey85", inherit.aes = FALSE) +
  geom_sf(data = coastlines, colour = "black", fill = NA, inherit.aes = FALSE) +
  coord_sf(xlim = c(-180,180), ylim = c(-90,90),
           expand = FALSE) +
  cowplot::theme_map() +
  theme(legend.position = c(0.75, 0.25),
        legend.justification = "center",
        legend.direction = "horizontal",
        strip.text.x = element_text(size = 12, colour = "black"))
```

## Extract SNR for biogeographic realms

Here we need to extract the SNR across all three scenarios within each biogeographic realm. The SNR values are then normalised to `{0, 1}` across time before the overlap within each realm is calculated.

```{r extractSNR, results='hide', eval = TRUE, message=FALSE}
# Data.table of Region numbers and corresponding name
reg_name <- data.table(
  Realm = c(1L:30L),
  Name = c("Inner Baltic Sea","Black Sea","NE Atlantic",
           "Norwegian Sea","Mediterranean","Artic Seas",
           "N Pacific","N American Boreal",
           "mid-tropical N Pacific","SE Pacific",
           "Caribbean & Gulf of Mexico","Gulf of California",
           "Indo-Pacific seas & Indian Ocean",
           "Gulfs of Aqaba, Aden, Suez, Red Sea","Tasman Sea","Coral Sea",
           "Mid South Tropical Pacific",
           "Offshore & NW North Atlantic","Offshore Indian Ocean",
           "Offshore W Pacific","Offshore S Atlantic",
           "Offshore mid-E Pacific","Gulf of Guinea",
           "Rio de La Plata","Chile","S Australia",
           "S Africa","New Zealand","N W Pacific",
           "Southern Ocean"))

# Merge region names
regions <- merge(regions, reg_name, by = "Realm")

realms <- unique(regions$Name)

# iterate through each realm and return a data.table of SNR values
reg_SNR <- lapply(realms, function(r) {
  sub <- regions[regions$Name == r, ]
  st_agr(sub) <- "constant"
  templ_ras <- rast(resolution = 2.5, crs = "EPSG:4326",
                    ## round bbox coord to 2.5°
                    xmin = round_any(st_bbox(sub)[1], 2.5), 
                    xmax = round_any(st_bbox(sub)[3], 2.5),
                    ymin = round_any(st_bbox(sub)[2], 2.5),
                    ymax = round_any(st_bbox(sub)[4], 2.5))
  sub <- rasterize(vect(sub), y = templ_ras, cover = TRUE)
  # only keep cells that have >= 15% cover
  sub[sub >= 0.15] <- 1
  sub[sub != 1] <- NA
  # Crop the raster data and mask
  snr_prj <- mask(x = crop(snr_comb, sub),
                  mask = sub)
  # Rescale values to {0, 1} across time
  sMax <- max(as.vector(global(snr_prj, "max", na.rm = TRUE)[,1]))
  sMin <- min(as.vector(global(snr_prj, "min", na.rm = TRUE)[,1]))
  snr_prj <- stretch(snr_prj, minv = 0, maxv = 1, smin = sMin, smax = sMax)
  # convert to data.table
  ext <- as.data.table(as.data.frame(snr_prj, xy = TRUE, na.rm = TRUE))
  ext <- melt(ext, id.vars = c("x", "y"), value.name = "SNR", variable.name = "Scenario")
  ext[, Region := r][, ID := 1:.N]
  setcolorder(ext, c("ID", "x", "y", "Region", "Scenario", "SNR"))
  return(ext)
})

# Covert list to data.table
reg_SNR <- rbindlist(reg_SNR)
setorder(reg_SNR, Region, ID)

# Number of samples within each region
nCells <- reg_SNR[, .(N = .N), by = c("Region", "Scenario")][Scenario == "TraCE", ]

# Export to CSV
fwrite(reg_SNR, "outputs/SNR_Cell.csv")
```

We can see that the SNR values are lower in the paleo period, than under the two RCP scenarios. Also, SNR under RCP 8.5 is higher than that under RCP 4.5

```{r SNRcomparison, echo = FALSE, fig.show='asis', fig.cap="SNR values within each region, standardised across time. n = number of cells in each region.", fig.align='center', fig.height=12, fig.width=12, out.width='100%'}
ggplot(reg_SNR, aes(x = Scenario, y = SNR,
                    group = interaction(Region, Scenario),
                    fill = Scenario)) +
  facet_wrap(~Region, ncol = 5, nrow = 6, scales = "fixed",
             labeller = label_wrap_gen(width = 20)) +
  geom_bar(position = "dodge", stat = "identity",
           show.legend = FALSE) +
  scale_fill_manual(values = c(
    "TraCE" = "#1b9e77",
    "RCP45" = "#7570b3",
    "RCP85" = "#d95f02")) +
  scale_y_continuous(breaks = scales::pretty_breaks(5)(0:1),
                     limits = c(0, 1),
                     expand = expansion(mult = c(0, .10))) +
  geom_text(data = nCells ,
            aes(x = "TraCE", y = 1.0, label = sprintf("n = %.0f", N)),
            colour = "black") +
  theme_half_open(font_size = 10) +
  theme(strip.background = element_blank(),
        strip.text = element_text(colour = "black", hjust = 0)) +
  panel_border() +
  background_grid()
```

From the sample sizes in the plot above we'll exclude the following regions: (i) Black Sea; (ii) Gulfs of Aqaba, Aden, Suez, Red Sea; and (iii) Inner Baltic Sea, from any further analysis.

## Histograms of overlap

Here we'll generate density plots of SNR values within each region which can then be used to estimate the overlap between SNR values in the past and future under both RCP scenarios.

```{r overlapHisto, echo = FALSE, fig.show='asis', fig.cap="Figure S2 - Density plots of SNR within each marine bioregion", fig.align='center', fig.height=12, fig.width=12, out.width='100%'}
exc_reg <- nCells[N < 5, ][["Region"]]
# Calculate median and quantiles for plotting
# Calculate Q10, 50, and 90 to add to plot
signif.lines <- reg_SNR[!Region %in% exc_reg, 
                        .(xMed = quantile(SNR, 0.5),
                          x10 = quantile(SNR, 0.10),
                          x90 = quantile(SNR, 0.90)),
                        by = c("Scenario", "Region")]

figs1 <- ggplot(data = reg_SNR[!Region %in% exc_reg, ],
                    aes(x = SNR, #y = ..scaled..,
                        colour = Scenario,
                    group = interaction(Region, Scenario),
                    fill = Scenario)) +
  facet_wrap(~Region, ncol = 5, nrow = 6, scales = "free_y",
             labeller = label_wrap_gen(width = 20)) +
  geom_density(alpha = 1/3, position = "identity", 
               size = 0.5, trim = TRUE,
               kernel = "epanechnikov", bw = "SJ",
               trim = TRUE, adjust = 1.5) +
  scale_fill_manual(values = c(
    "TraCE" = "#2166ac",
    "RCP45" = "#f4a582",
    "RCP85" = "#b2182b"),
    guide = "legend") +
  scale_colour_manual(values = c(
    "TraCE" = "#2166ac",
    "RCP45" = "#f4a582",
    "RCP85" = "#b2182b")) +
  scale_x_continuous(limits = c(0,1)) +
  scale_y_continuous() +
  geom_vline(data = signif.lines,
             aes(xintercept = xMed, colour = Scenario,
                 group = interaction(Region, Scenario)), size = 0.5,
             show.legend = FALSE) +
  theme_half_open(font_size = 10) +
  panel_border() +
  theme(legend.position = c(0.70, 0.05),
        legend.direction = "horizontal",
        legend.justification = "center",
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(colour = "black", hjust = 0),
        panel.spacing = unit(0.5, "lines")) +
  guides(fill = guide_legend(title.position = "top",
                             title.hjust = 0.5,
                             label.position = "bottom",
                             label.hjust = 0.5,
                             keywidth = unit(3, "mm"),
                             keyheight = unit(3, "mm"),
                             nrow = 1),
         colour = guide_legend(title.position = "top",
                             title.hjust = 0.5,
                             label.position = "bottom",
                             label.hjust = 0.5,
                             keywidth = unit(3, "mm"),
                             keyheight = unit(3, "mm"),
                             nrow = 1)) +
  labs(x = NULL)
ggsave(filename = "FigureS1.png", plot = figs1,
       device = "png", type = "cairo",
       path = "Figures",
       width = 12, height = 12, units = "in",
       dpi = 320, bg = "white")
figs1
```

## Calculate overlap

Here we'll calculate the overlap between past and future SNR for each region with each scenario.

```{r calcOverlap, results='markup'}
realms <- unique(regions$Name)
# exclude the small realms
realms <- realms[!realms %in% nCells[N < 5, ][["Region"]]]

regional_overlap <- rbindlist(lapply(realms, function(r) {
  sub <- reg_SNR[Region == r, ]
  fut_DT <- rbindlist(lapply(c("RCP45", "RCP85"), function(future) {
    rcp_SNR <- sub[Scenario %in% c("TraCE", future), ]
    # paleo
    dens_a <- rcp_SNR[Scenario == "TraCE", ][["SNR"]]
    # RCP
    dens_b <- rcp_SNR[Scenario != "TraCE", ][["SNR"]]
    # Overlap
    kde_over <- overlap(x = list(trace = dens_a, rcp = dens_b),
                        boundaries = list(
                          from = c(min(dens_a), min(dens_b)),
                          to = c(max(dens_a), max(dens_b))
                        ), kernel = "epa",
                        nbins = 100, plot = FALSE,
                        cut = TRUE, bw = "SJ",
                        adjust = 1.5)
    # Kolmogorov-Smirnov Test
    ## Test is sensitive to changes in shape, spread, or median.
    ## Substantial differences in the above will result in a small P value
    kde_KS <- ks.test(x = dens_a, y = dens_b, exact = TRUE)
    # Stars for significance
    stars <- stars.pval(kde_KS$p.value)
    fut <- data.table(Region = r, Scenario = future,
                         Overlap = round(kde_over$OV,2),
                         KS = kde_KS$statistic,
                         KS_P = kde_KS$p.value, 
                         KS_S = stars)
    return(fut)
  }), use.names = TRUE)
  return(fut_DT)
}), use.names = TRUE)

## adjust p-values
regional_overlap[, KS_Padj := p.adjust(KS_P, method = "BY")
                 ][, 
                   KS_Sadj := gtools::stars.pval(KS_Padj)]

# Save results to CSV
fwrite(regional_overlap, "outputs/TableS1.csv")

# summary by scenario
regional_overlap[, .(Max = max(Overlap),
                     Min = min(Overlap),
                     Mean = round(mean(Overlap), 2)),
      by = "Scenario"]

# summary by region
regional_overlap[, .(Max = max(Overlap),
                     Min = min(Overlap),
                     Mean = round(mean(Overlap), 2)),
      by = "Region"][order(-Max), ]

```

## Overlap raster

Here we'll create a raster of overlap for both scenarios

```{r overlapRaster, results='markup'}
# 0.5° template raster for plotting
templ_ras <- rast(res = 0.5, extent = ext(rast()), crs = "EPSG:4326")

# Join RCP 4.5 overlap
regions <- merge(regions, regional_overlap[Scenario == "RCP45", c("Region", "Overlap")],
                 by.x = "Name", by.y = "Region")
colnames(regions)[5] <- "RCP45_Overlap"

# Join RCP 8.5 overlap
regions <- merge(regions, regional_overlap[Scenario == "RCP85", c("Region", "Overlap")],
                 by.x = "Name", by.y = "Region")
colnames(regions)[6] <- "RCP85_Overlap"
st_write(regions, "outputs/output_shapes.gpkg",
         layer = "marine_regions_SNR_overlap", quiet = TRUE)

# Convert to raster
regions_overlap45 <- rasterize(vect(regions), y = templ_ras, 
                               field = c("RCP45_Overlap"))
regions_overlap85 <- rasterize(vect(regions), y = templ_ras, 
                               field = c("RCP85_Overlap"))
regions_overlap <- rast(list(regions_overlap45, regions_overlap85))

regions_overlap

writeRaster(regions_overlap, 
            "outputs/marine_regions_SNR_overlap.grd")
```

A quick plot of the overlap in each of the regions shows the following

```{r plotOV, echo = FALSE, fig.show='asis', fig.cap="Figure S3 - Plot of SNR overlap within each marine bioregion", fig.align='center', fig.height = 6.5, fig.width = 5, out.width='100%'}
xdf <- setDT(as.data.frame(regions_overlap, xy = TRUE, na.rm = TRUE))
xdf <- melt(xdf, id.vars = c("x", "y"), variable.name = "Scenario",
            value.name = "Overlap")
ggplot() +
  geom_raster(data = xdf, aes(x, y, fill = Overlap)) +
  facet_wrap(~Scenario, ncol = 1,
             labeller = as_labeller(c("RCP45_Overlap" = "RCP 4.5",
                                  "RCP85_Overlap" = "RCP 8.5"))) +
  scale_fill_stepsn(colours = hcl.colors(17, "Temps", rev = TRUE),
                       limits = c(0, 0.80),                       
                       breaks = scales::pretty_breaks(16)(seq(0, 0.80, length = 8)),
                       labels = function(x) ifelse(round((x *100) %% 2) == 0, x*100 , ""),
                       oob = scales::squish,
                       guide = guide_colourbar(
                         title = "Overlap (%)",
                         title.position = "top",
                         title.hjust = 0.5,
                         title.theme = element_text(size = 12, colour = "black"),
                         label.position = "bottom",
                         label.hjust = 0.5,
                         label.theme = element_text(size = 12, colour = "black"),
                         barwidth = 12,
                         barheight = 0.5,
                         nbin = 100,
                         frame.colour = "black",
                         frame.linewidth = 0.5
                       )) +
  geom_sf(data = regions, inherit.aes = FALSE, fill = NA, colour = "grey40") +
  geom_sf(data = coastlines, colour = "black", inherit.aes = FALSE) +
  coord_sf(xlim = c(-180,180), ylim = c(-90,90),
           expand = FALSE) +
  theme_map() +
  theme(legend.position = "bottom",
        legend.justification = "center",
        legend.direction = "horizontal",
        strip.text.x = element_text(size = 12, colour = "black"))
```

## Overlap summaries

What is the global weighted overlap under both scenarios?

```{r weightedOverlap, results='markup'}
# global weighted overlap under both scenarios
sapply(c("RCP45_Overlap", "RCP85_Overlap"), function(i) {
  round(wtd.mean(x = regions[[i]]*100,
                weights = as.vector(regions$Area),
                normwt = FALSE), 1)
})
```

## Figure 2 and Figure 3

Here we make figure 2 and figure 3. Figure 2 is made using the overlap data for RCP 4.5, but can be easily changed to used RCP 8.5.

```{r makeFig2, echo = FALSE, fig.width=9, out.width='100%', results='markup', fig.show='hold'}
mollPrj <- 'PROJCS["Pacific_Centrered_Mollweide",
 GEOGCS["GCS_WGS_1984",
  DATUM["D_WGS_1984",
   SPHEROID["WGS_1984",6378137.0,298.257223563]],
  PRIMEM["Greenwich",0.0],
  UNIT["Degree",0.0174532925199433]],
 PROJECTION["Mollweide"],
 PARAMETER["False_Easting",0.0],
 PARAMETER["False_Northing",0.0],
 PARAMETER["Central_Meridian",160],
 UNIT["Meter",1.0]]'

templ_ras <- rast(resolution = 2.5, crs = "EPSG:4326")
templ_ras <- project(templ_ras, y = mollPrj)
res(templ_ras) <- c(50000) # 50km only for plotting

# Projected overlap map
ovrProj <- project(regions_overlap[["RCP45_Overlap"]], 
                   y = templ_ras, method = "near")
ovrProj <- setDT(as.data.frame(ovrProj, xy = TRUE))

# projected shapefiles for aesthetics
hotspots_prj <- read_sf("outputs/output_shapes.gpkg", layer = "marine_hotspots")
hotspots_prj <- hotspots_prj[hotspots_prj$Thresh == 95, ]
hotspots_prj <- rotate_prj(hotspots_prj, crs = st_crs(templ_ras))
land_prj <- read_sf("shapes/mapping_shapes.gpkg", layer = "land")
regions_prj <- read_sf("shapes/mapping_shapes.gpkg", layer = "regions_merged")
coast_prj <- read_sf("shapes/mapping_shapes.gpkg", layer = "coast")
bbox_prj <- read_sf("shapes/mapping_shapes.gpkg", layer = "bbox")

# fig 2 map
map02 <- ggplot() +
  geom_raster(data = ovrProj, aes(x = x, y = y, fill = RCP45_Overlap),
              show.legend = TRUE) +
  geom_sf(data = land_prj, size = 0.25,
          inherit.aes = FALSE, colour = NA, fill = "#E9E9E9") +
  geom_sf(data = regions_prj, size = 0.5,
          inherit.aes = FALSE, fill = NA, colour = "grey40") +
   geom_sf_pattern(data = hotspots_prj, size = 0.5,
                   pattern = "crosshatch", pattern_size = 0.25,
                   pattern_density = 0.05, pattern_spacing = 0.01,
                   pattern_colour  = "white",
                   inherit.aes = FALSE, colour = "white", fill = NA) +
  geom_sf(data = coast_prj, size = 0.25,
          inherit.aes = FALSE, fill = "#E9E9E9", colour = "black") +
  geom_sf(data = bbox_prj, size = 0.25,
          inherit.aes = FALSE, fill = NA, colour = "black") +
  scale_fill_stepsn(colours = cptcity::cpt("jjg_cbcont_div_cbcSpectral", 20)[-c(18:20)],
                    breaks = seq(0, 0.80, length = 17),
                    limits = c(-0.001, 0.80),
                    labels = function(x) ifelse(round((x *100)) %% 2L == 0, x*100 , ""),
                    oob = scales::squish,
                    na.value = "transparent",
                    guide = guide_colourbar(
                         title = "Overlap (%)",
                         title.position = "top",
                         title.hjust = 0.5,
                         title.theme = element_text(size = 12, colour = "black"),
                         label.position = "bottom",
                         label.hjust = 0.5,
                         label.theme = element_text(size = 12, colour = "black"),
                         barwidth = 12,
                         barheight = 1,
                         nbin = 100,
                         frame.colour = "black",
                         frame.linewidth = 0.5,
                         ticks = TRUE,
                         ticks.colour = "black",
                         ticks.linewidth = 0.5,
                         draw.ulim = FALSE,
                         draw.llim = FALSE
                       )) +
  coord_sf(expand = FALSE) +
  theme_map() +
  theme(legend.position = c(0.50, -0.125),
        legend.justification = "center",
        legend.direction = "horizontal",
        strip.text.x = element_text(size = 12, colour = "black"))

# Density plots for borders
plot_regions <- c("N American Boreal", "Caribbean & Gulf of Mexico",
                   "Gulf of California", "Coral Sea")

signif.lines <- reg_SNR[Region %in% plot_regions,
                        .(xMed = quantile(SNR, 0.5),
                          x5 = quantile(SNR, 0.10),
                          x95 = quantile(SNR, 0.95)),
                        by = c("Scenario", "Region")]

plot_list <- lapply(plot_regions, function(region) {
  reg_dens <- ggplot(data = reg_SNR[Region == region & Scenario != "RCP85", ],
         aes(x = SNR, colour = Scenario,
             group = interaction(Region, Scenario),
             fill = Scenario)) +
  geom_density(alpha = 1/4, position = "identity", 
               aes(y = ..scaled..),
               size = 0.5, kernel = "epanechnikov", bw = "SJ",
               trim = TRUE, adjust = 1.5) +
  scale_fill_manual(values = c(
    "TraCE" = "#2166ac",
    "RCP45" = "#f4a582",
    "RCP85" = "#b2182b"),
    guide = "legend") +
  scale_colour_manual(values = c(
    "TraCE" = "#2166ac",
    "RCP45" = "#f4a582",
    "RCP85" = "#b2182b")) +
  scale_x_continuous(limits = c(0,1), expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
   geom_segment(data = signif.lines[Region == region & Scenario != "RCP85", ],
                 aes(y = 0.01, yend = 0.33, x = xMed, xend = xMed,
                     colour = Scenario, group = interaction(Region, Scenario)), 
                     arrow = grid::arrow(ends = "first", angle = 30, length = unit(.25,"cm"))) +
  theme_void() +
    theme(legend.position = "none")
  return(reg_dens)
})

names(plot_list) <- plot_regions

left_col <- plot_list$`N American Boreal`/plot_spacer()/plot_list$`Coral Sea`/plot_spacer()
right_col <- plot_list$`Caribbean & Gulf of Mexico`/plot_spacer()/plot_list$`Gulf of California`/plot_spacer()

# Read in the images for overlay
walrus <- image_read("Figures/Stock/walrus_small.jpg")
manatee <- image_read("Figures/Stock/manatee_small.jpg")
reef <- image_read("Figures/Stock/reef_small.jpg")
hammer <- image_read("Figures/Stock/hammerhead_small.jpg")

fig02 <- ggdraw({{left_col} | {map02} | {right_col}} + 
  plot_layout(ncol = 3, widths = c(0.25,1,0.25))) +
  draw_label(label = "a", size = 16, x = 0.25, y = 0.75, fontface = "bold") +
  ## Coral sea
  draw_line(x = c(0.17, 0.43),
            y = c(0.43, 0.43),
            size = 0.65,
            arrow = grid::arrow(ends = "last", angle = 30, length = unit(.25,"cm"))) +
  draw_label(label = "d", size = 16, x = 0.025, y = 0.45, fontface = "bold") +
  ## NA Boreal
  draw_line(x = c(0.17, 0.35, 0.35),
            y = c(0.85, 0.85, 0.78),
           size = 0.65,
            arrow = grid::arrow(ends = "last", angle = 30, length = unit(.25,"cm"))) +
  draw_label(label = "b", size = 16, x = 0.025, y = 0.95, fontface = "bold") +
  ## Gulf of Mexico
  draw_line(x = c(0.825, 0.70, 0.70),
            y = c(0.85, 0.85, 0.62),
            size = 0.65,
            arrow = grid::arrow(ends = "last", angle = 30, length = unit(.25,"cm"))) +
   draw_label(label = "c", size = 16, x = 0.835, y = 0.95, fontface = "bold") +
  ## Gulf of Cali.
  draw_line(x = c(0.825, 0.70, 0.70),
            y = c(0.43, 0.43, 0.51),
            size = 0.65,
            arrow = grid::arrow(ends = "last", angle = 30, length = unit(.25,"cm"))) +
  draw_label(label = "e", size = 16, x = 0.835, y = 0.45, fontface = "bold")
fig02 <- ggdraw(fig02) +
  ## reef
  draw_image(reef, x = 0.0125, y = -0.345, clip = "off", width = 0.16) +
  ## walrus
  draw_image(walrus, x = 0.0125, y = 0.125, clip = "off", width = 0.16) +
  ## manatee
  draw_image(manatee, x = 0.825, y = 0.125, clip = "off", width = 0.16) +
  ## hammerheads
  draw_image(hammer, x = 0.825, y = -0.345, clip = "off", width = 0.16)

# Read in the temperature data for the hotspots
hs_temps <- setDT(readRDS("Source/annual_temps_latbands_hotspots_v2.RDS"))

trace <- ggplot(data = hs_temps[Scen == "TraCE" & value >= 18 & Year < -100, ], 
                aes(x = Year/1000, y = lyr1, fill = value)) +
  geom_raster(na.rm = TRUE) +
  facet_wrap(~Scen, ncol = 1, labeller = as_labeller(c("TraCE" = "a"))) +
  scale_fill_stepsn(colours = cpt("arendal_temperature", n = 24)[-(1:9)],
                    limits = c(18, 32),
                    oob = scales::squish,
                    na.value = "transparent",
                    breaks = seq(18, 32, length = 15),
                    labels = function(x) ifelse(x %% 2L == 0, x, "")) +
  scale_x_continuous(breaks = seq(-21,0,length = 7), expand = c(0, 0),
                     limits = c(-21,0)) +
  scale_y_continuous(limits = c(-45, 40), 
                     expand = c(0,0),
                     breaks = seq(-40,40,20)) +
  geom_vline(xintercept = c(-9, -5),
             colour = "#000000", size = 0.5, linetype = "dotted") +
  theme_half_open() +
  theme(legend.position = "none",
        legend.direction = "horizontal",
        legend.justification = "centre",
        strip.text.x = element_text(size = 14, colour = "black", face = "bold", hjust = 0),
        strip.background = element_blank(),
        plot.background = element_rect(fill = NA, colour = NA),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        axis.line = element_line(size = 0.25, colour = "black"),
        axis.ticks.x = element_line(size = 0.25, colour = "black"),
        axis.ticks.y = element_line(size = 0.25, colour = "black"),
        plot.margin = unit(c(0,0.5,-0.5,0), "lines")) +
  labs(y = "Latitude (°)", x = "Year (ka BP)")

rcp <- ggplot(data = hs_temps[Scen %in% c("RCP45", "RCP85") & value >= 18, ], 
                aes(x = Year, y = lyr1, fill = value)) +
  facet_wrap(~Scen, ncol = 2, 
             labeller = as_labeller(c("RCP45" = "b", "RCP85" = "c"))) +
  geom_raster(na.rm = TRUE) +
  scale_fill_stepsn(colours = cpt("arendal_temperature", n = 24)[-(1:9)],
                    limits = c(18, 32),
                    oob = scales::squish,
                    na.value = "transparent",
                    breaks = seq(18, 32, length = 15),
                    labels = function(x) ifelse(x %% 2L == 0, x, ""),
                       guide = guide_colourbar(
                         title = "Temperature (°C)",
                         title.position = "top",
                         title.hjust = 0.5,
                         title.theme = element_text(size = 10, colour = "black"),
                         label.position = "bottom",
                         label.hjust = 0.5,
                         label.theme = element_text(size = 10, colour = "black"),
                         barwidth = 10,
                         barheight = 0.75,
                         nbin = 100,
                         frame.colour = "black",
                         frame.linewidth = 0.5,
                         ticks = TRUE,
                         ticks.colour = "black",
                         ticks.linewidth = 0.5,
                         draw.ulim = FALSE,
                         draw.llim = FALSE
                       )) +
  geom_vline(xintercept = c(2080, 2100),
             colour = "#000000", size = 0.5, linetype = "dotted") +
  scale_x_continuous(breaks = seq(1850,2100, length = 6), 
                     expand = c(0, 0)) +
  scale_y_continuous(limits = c(-45, 40), 
                     expand = c(0,0),
                     breaks = seq(-40,40,20)) +
  coord_cartesian(clip = "off") +
  theme_half_open() +
  theme(legend.position = "bottom",
        legend.direction = "horizontal",
        legend.justification = "centre",
        legend.margin = margin(-0.5, 0, 0, 0, unit = "lines"),
        strip.text.x = element_text(size = 14, colour = "black", face = "bold", hjust = 0),
        strip.background = element_blank(),
        plot.background = element_rect(fill = NA, colour = NA),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        axis.line = element_line(size = 0.25, colour = "black"),
        axis.ticks.x = element_line(size = 0.25, colour = "black"),
        axis.ticks.y = element_line(size = 0.25, colour = "black"),
        panel.spacing.x = unit(1.75, "lines"),
        plot.margin = unit(c(0,0.5,-0.5,0), "lines")) +
  labs(y = "Latitude (°)", x = "Year (C.E.)")

fig03 <- {trace}/{rcp} + plot_layout(ncol = 1)

# save to PDF
ggsave(filename = "Figure02.pdf", plot = fig02,
       device = cairo_pdf, path = "Figures",
       width = 9, height = 4.2, units = "in", dpi = 350)
ggsave(filename = "Figure03.pdf", plot = fig03,
       device = cairo_pdf, path = "Figures",
       width = 9, height = 5, units = "in", dpi = 350)

# save to PNG for MS
ggsave(filename = "Figure02.png", plot = fig02,
       device = "png", type = "cairo", path = "Figures",
       bg = "white", width = 9, height = 4.2, units = "in", dpi = 350)
ggsave(filename = "Figure03.png", plot = fig03,
       device = "png", type = "cairo", path = "Figures",
       bg = "white", width = 9, height = 5, units = "in", dpi = 350)
# plot
fig02 + theme_map() + labs(caption = "Figure 02 - Exposure to future changes in rates of accelerated oceanic warming under RCP 4.5.")
fig03 + labs(caption = "Figure 03 - Temporal changes in latitudinal mean annual sea-surface temperature in hotspots of marine biodiversity")
```

Here we make Figure S2 - the overlap plot for RCP 8.5

```{r figS2, echo = FALSE, fig.width=9, out.width='100%', results='hide', fig.show='hold'}
# Projected bivariate map
ovrProj <- project(regions_overlap[["RCP85_Overlap"]], 
                   y = templ_ras, method = "near")
ovrProj <- setDT(as.data.frame(ovrProj, xy = TRUE))

# fig S2 map
maps2 <- ggplot() +
  geom_raster(data = ovrProj, aes(x = x, y = y, fill = RCP85_Overlap),
              show.legend = TRUE) +
  geom_sf(data = land_prj, size = 0.25,
          inherit.aes = FALSE, colour = NA, fill = "#E9E9E9") +
  geom_sf(data = regions_prj, size = 0.5,
          inherit.aes = FALSE, fill = NA, colour = "grey40") +
   geom_sf_pattern(data = hotspots_prj, size = 0.5,
                   pattern = "crosshatch", pattern_size = 0.25,
                   pattern_density = 0.05, pattern_spacing = 0.01,
                   pattern_colour  = "white",
                   inherit.aes = FALSE, colour = "white", fill = NA) +
  geom_sf(data = coast_prj, size = 0.25,
          inherit.aes = FALSE, fill = "#E9E9E9", colour = "black") +
  geom_sf(data = bbox_prj, size = 0.25,
          inherit.aes = FALSE, fill = NA, colour = "black") +
  scale_fill_stepsn(colours = cptcity::cpt("jjg_cbcont_div_cbcSpectral", 20)[-c(18:20)],
                    breaks = seq(0, 0.80, length = 17),
                    limits = c(-0.001, 0.80),
                    labels = function(x) ifelse(round((x *100)) %% 2L == 0, x*100 , ""),
                    oob = scales::squish,
                    na.value = "transparent",
                    guide = guide_colourbar(
                         title = "Overlap (%)",
                         title.position = "top",
                         title.hjust = 0.5,
                         title.theme = element_text(size = 12, colour = "black"),
                         label.position = "bottom",
                         label.hjust = 0.5,
                         label.theme = element_text(size = 12, colour = "black"),
                         barwidth = 12,
                         barheight = 1,
                         nbin = 100,
                         frame.colour = "black",
                         frame.linewidth = 0.5,
                         ticks = TRUE,
                         ticks.colour = "black",
                         ticks.linewidth = 0.5,
                         draw.ulim = FALSE,
                         draw.llim = FALSE
                       )) +
  coord_sf(expand = FALSE) +
  theme_map() +
  theme(legend.position = c(0.50, -0.125),
        legend.justification = "center",
        legend.direction = "horizontal",
        strip.text.x = element_text(size = 12, colour = "black"))

signif.lines <- reg_SNR[Region %in% plot_regions,
                        .(xMed = quantile(SNR, 0.5),
                          x5 = quantile(SNR, 0.10),
                          x95 = quantile(SNR, 0.95)),
                        by = c("Scenario", "Region")]

plot_list <- lapply(plot_regions, function(region) {
  reg_dens <- ggplot(data = reg_SNR[Region == region & Scenario != "RCP45", ],
         aes(x = SNR, colour = Scenario,
             group = interaction(Region, Scenario),
             fill = Scenario)) +
  geom_density(alpha = 1/4, position = "identity", 
               aes(y = ..scaled..),
               size = 0.5, trim = TRUE,
               kernel = "epanechnikov", bw = "SJ",
               trim = TRUE, adjust = 1.5) +
  scale_fill_manual(values = c(
    "TraCE" = "#2166ac",
    "RCP45" = "#f4a582",
    "RCP85" = "#b2182b"),
    guide = "legend") +
  scale_colour_manual(values = c(
    "TraCE" = "#2166ac",
    "RCP45" = "#f4a582",
    "RCP85" = "#b2182b")) +
  scale_x_continuous(limits = c(0,1), expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
    geom_segment(data = signif.lines[Region == region & Scenario != "RCP45", ],
                 aes(y = 0.01, yend = 0.33, x = xMed, xend = xMed,
                     colour = Scenario, group = interaction(Region, Scenario)), 
                 arrow = grid::arrow(ends = "first", angle = 30, length = unit(.25,"cm"))) +
    theme_void() +
    theme(legend.position = "none")
  return(reg_dens)
})

names(plot_list) <- plot_regions

left_col <- plot_list$`N American Boreal`/plot_spacer()/plot_list$`Coral Sea`/plot_spacer()
right_col <- plot_list$`Caribbean & Gulf of Mexico`/plot_spacer()/plot_list$`Gulf of California`/plot_spacer()

figs2 <- ggdraw({{left_col} | {maps2} | {right_col}} + 
  plot_layout(ncol = 3, widths = c(0.25,1,0.25))) +
  draw_label(label = "a", size = 16, x = 0.25, y = 0.75, fontface = "bold") +
  ## Coral sea
  draw_line(x = c(0.17, 0.43),
            y = c(0.43, 0.43),
            size = 0.65,
            arrow = grid::arrow(ends = "last", angle = 30, length = unit(.25,"cm"))) +
  draw_label(label = "d", size = 16, x = 0.025, y = 0.45, fontface = "bold") +
  ## NA Boreal
  draw_line(x = c(0.17, 0.35, 0.35),
            y = c(0.85, 0.85, 0.78),
           size = 0.65,
            arrow = grid::arrow(ends = "last", angle = 30, length = unit(.25,"cm"))) +
  draw_label(label = "b", size = 16, x = 0.025, y = 0.95, fontface = "bold") +
  ## Gulf of Mexico
  draw_line(x = c(0.825, 0.70, 0.70),
            y = c(0.85, 0.85, 0.62),
            size = 0.65,
            arrow = grid::arrow(ends = "last", angle = 30, length = unit(.25,"cm"))) +
   draw_label(label = "c", size = 16, x = 0.835, y = 0.95, fontface = "bold") +
  ## Gulf of Cali.
  draw_line(x = c(0.825, 0.70, 0.70),
            y = c(0.43, 0.43, 0.51),
            size = 0.65,
            arrow = grid::arrow(ends = "last", angle = 30, length = unit(.25,"cm"))) +
  draw_label(label = "e", size = 16, x = 0.835, y = 0.45, fontface = "bold")

figs2 <- ggdraw(figs2) +
  ## reef
  draw_image(reef, x = 0.0125, y = -0.345, clip = "off", width = 0.16) +
  ## walrus
  draw_image(walrus, x = 0.0125, y = 0.125, clip = "off", width = 0.16) +
  ## manatee
  draw_image(manatee, x = 0.825, y = 0.125, clip = "off", width = 0.16) +
  ## hammerheads
  draw_image(hammer, x = 0.825, y = -0.345, clip = "off", width = 0.16)


# save to PDF
ggsave(filename = "FigureS2.pdf", plot = figs2,
       device = cairo_pdf, path = "Figures",
       width = 9, height = 4.2, units = "in", dpi = 350)

# save to PNG for MS
ggsave(filename = "FigureS2.png", plot = figs2,
       device = "png", type = "cairo", path = "Figures",
       bg = "white", width = 9, height = 4.2, units = "in", dpi = 350)

# plot
figs2 + theme_map() + labs(caption = "Figure S2 - Exposure to future changes in rates of accelerated oceanic warming under RCP 8.5.")
```

