---
title: "05 - Hotspot Delta SNR"
author: "Stuart Brown"
output: 
  html_document: 
    theme: flatly
    toc: yes
    toc_depth: 3
    toc_float: true
    fig_width: 7
    fig_height: 5
    fig_caption: yes
    number_sections: no
    keep_md: yes
    df_print: paged
    code_folding: show
    highlight: tango
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, rows.print = 10, cols.print = 10,
                      rownames.print = FALSE, warning = FALSE,
                      message = FALSE, eval = TRUE, results = "markup", 
                      error = FALSE, strip.white = TRUE,
                      tidy = "styler", dpi = 320, fig.align = "center",
                      strip.white = TRUE, 
                      dev.args = list(png  = list(type = "cairo")))
```

## Read in the data

Here we read in the climate data and species richness data for plotting

```{r importData, results='markup'}
library(terra)
library(sf)
library(units)
library(ggplot2)
library(data.table)
library(rnaturalearth)

# Ocean
coastlines <- ne_coastline(50, returnclass = "sf")
oceans <- read_sf("outputs/output_shapes.gpkg", layer = "ne_ocean")

# Marine realms from Costello et al
regions <- read_sf("outputs/output_shapes.gpkg", layer = "marine_regions_clipped")

st_agr(oceans) <- "constant"
st_agr(regions) <- "constant"

# Area of each region in km^2
regions$Area <- set_units(st_area(regions), "km^2")

# Region names
reg_name <- data.table(
  Realm = c(1L:30L),
  Name = c("Inner Baltic Sea","Black Sea","NE Atlantic",
           "Norwegian Sea","Mediterranean","Artic Seas",
           "N Pacific","N American Boreal",
           "mid-tropical N Pacific","SE Pacific",
           "Caribbean & Gulf of Mexico","Gulf of California",
           "Indo-Pacific seas & Indian Ocean",
           "Gulfs of Aqaba, Aden, Suez, Red Sea","Tasman Sea","Coral Sea",
           "Mid South Tropical Pacific",
           "Offshore & NW North Atlantic","Offshore Indian Ocean",
           "Offshore W Pacific","Offshore S Atlantic",
           "Offshore mid-E Pacific","Gulf of Guinea",
           "Rio de La Plata","Chile","S Australia",
           "S Africa","New Zealand","N W Pacific",
           "Southern Ocean"))
regions <- merge(regions, reg_name, by = "Realm")

# 95th percentile hotspots
hotspots <- read_sf("outputs/output_shapes.gpkg", layer = "marine_hotspots")
hotspots <- hotspots[hotspots$Thresh == 95, ]
st_agr(hotspots) <- "constant"
```

## Intersect hotspots with marine biogeographic realms

Here we need to intersect the hotspots with their corresponding biogeographic realm

```{r intersectHSAndRealms}
hs_int <- st_transform(
  st_intersection(
    st_buffer(st_transform(hotspots, crs = "EPSG:32662"), 0),
    st_buffer(st_transform(regions[4], crs = "EPSG:32662"), 0)
    ), "EPSG:4326")
st_agr(hs_int) <- "constant"

# Area of intersected hotspots in km^2
hs_int$Area <- set_units(st_area(hs_int), "km^2")

# Area of 2.5° grid cells at 0° lat == 77270.818 km2
## remove intersected hotspots that have less than 3 cells of area
hs_int <- hs_int[hs_int$Area >= as_units(231813, "km^2"), ]

# Keep only the regions that contain hotspots
regions_w_hs <- regions[regions$Name %in% hs_int$Name, ]
```

A quick plot of the regions that contain hotspots, and the hotspot locations shows the following

```{r plotRegWHot, echo = TRUE, fig.show='asis', fig.cap="Plot of marine biogeographic realms the contain marine biodiversity hotspots", fig.align='center', fig.height = 5, fig.width = 8, out.width='100%'}
ggplot() + 
  geom_sf(data = oceans, fill = "grey90", colour = NA) +
  geom_sf(data = regions_w_hs, aes(fill = Name), colour = "white", size = 0.75) +
  scale_fill_manual(values = {set.seed(18);sample(hcl.colors(12, "Batlow"), 12, replace = FALSE)}) +
  geom_sf(data = hs_int, fill = NA, colour = "black") +
  geom_sf(data = coastlines, colour = "black", inherit.aes = FALSE, size = 0.5) +
  coord_sf(xlim = c(-180,180), ylim = c(-90,90),
           expand = FALSE) +
  cowplot::theme_map() +
  theme(legend.position = "bottom",
        legend.justification = "center",
        legend.direction = "horizontal",
        legend.text = element_text(size = 8, colour = "black"),
        legend.title = element_blank())
```

## Intersect hotspot locations with SNR

Now we need to intersect the hotspot locations with the SNR data so we can look at changes in SNR within each hotspot and biogeographic realm.

```{r hsIntersectSNR}
# Convert hotspots to points
## rasterise
hotspots_r <- rasterize(vect(hs_int),
                        y = rast(resolution = 2.5, crs = "EPSG:4326", 
                                 extent = c(-180,180,-90,90)),
                        field = "Name")
## convert to points
hotspots_p <- st_as_sf(as.points(hotspots_r))

# Read in the grid-cell SNR data
cell_snr <- fread("outputs/SNR_Cell.csv")
cell_snr <- cell_snr[Region %in% hotspots_p$Name, ]
cell_snr <- st_as_sf(cell_snr, coords = c("x", "y"), crs = 4326)

# intersect hotspots with cell_snr
## only intersect Region, Scenario, and SNR
hotspot_delta_snr <- st_transform(
  st_intersection(
    st_transform(hotspots_p, crs = "EPSG:32662"),
    st_transform(cell_snr[, c(2:4)], crs = "EPSG:32662")
    ), crs = "EPSG:4326")

# change to data.table
hotspot_delta <- as.data.table(hotspot_delta_snr)
hotspot_delta[, ':='(
  Lon = st_coordinates(hotspot_delta_snr)[, 1],
  Lat = st_coordinates(hotspot_delta_snr)[, 2],
  geometry = NULL)]
colnames(hotspot_delta)[1:2] <- c("HotspotRegion", "SNRCellRegion")
setorder(hotspot_delta, -Scenario, HotspotRegion)
hotspot_delta[1:20,]
```

There are a number of cells that cross boundaries between the regions.
These can be identified by rows where the cells have intersected, but the `Region` and `Name` columns have different values.
There are three scenarios so there should have been `r 3*nrow(hotspots_p)` rows returned in the intersection if all hotspots matched to an individual cell.
However, there are `r nrow(hotspot_delta)` rows returned, indicating that there are `r nrow(hotspot_delta) - (3*nrow(hotspots_p))` extra rows.
That means there are `r (nrow(hotspot_delta) - (3*nrow(hotspots_p)))/3` rows extra for each scenario.
Here's a preview of these rows for the RCP 4.5 and RCP 8.5 scenario.

```{r prevExtraRows, echo = FALSE, results='markup', fig.show='hold'}
hotspot_delta[(HotspotRegion != SNRCellRegion) & Scenario != "TraCE", ][order(HotspotRegion), ]
```

We're going to leave the extra rows in because there is no reason why hotspots that exist on the boundaries between two biogeographic realms shouldn't cross the boundaries.

## Extreme conditions

Here we test whether conditions in the future are more extreme than those seen in the past.
We do this by calculating the number of cells within a biogeographic realm and hotspots within that realm, that have higher SNR values than those experienced in the past.
We do this using the 95th percentile of past SNR within each realm.
A sensitivity analysis is also conducted $\pm$ 5% (`{90-100th percentile}`).

```{r extremeCondAssessment, results='markup', rows.print = 13, cols.print = 12}
# Read in the grid-cell SNR data agaain
cell_snr <- fread("outputs/SNR_Cell.csv")
cell_snr <- cell_snr[Region %in% hotspots_p$Name, ]

# quantiles from the past for each realm
qSNR <- setDT(cell_snr)[Scenario == "TraCE",
                        as.list(quantile(SNR, prob = seq(0.9,1,0.01))),
                        by = c("Region")]
qSNR

# Regions to iterate through
regions <- qSNR$Region

num_ext_cells <- lapply(regions, function(realm) {
  ## Region subset. This will include hotspots that cross the boundary
  sub <- copy(hotspot_delta[Scenario != "TraCE" & SNRCellRegion == realm, ])
  ## loop through the RCP scenarios
  realm_rcp <- lapply(unique(sub$Scenario), function(rcp) {
    sub_rcp <- sub[Scenario == rcp, ]
    ## all cells in a realm in the past
    realm_cells <- setDT(copy(cell_snr))
    realm_cells <- realm_cells[Scenario == "TraCE" & Region == realm, ]
    ## how many hotspots cells >= different quantiles
    ncell <- sapply(colnames(qSNR)[-1], function(thresh) {
     nrow(sub_rcp[SNR >= qSNR[Region == realm,][[thresh]], ]) 
    }, USE.NAMES = TRUE)
    sub_realm <- copy(cell_snr[Scenario == rcp & Region == realm, ])
    ncell_realm <- sapply(colnames(qSNR)[-1], function(thresh) {
     nrow(sub_realm[SNR >= qSNR[Region == realm,][[thresh]], ]) 
    }, USE.NAMES = TRUE)
    # percentage of extreme cells (ncell extreme hotspots/total number of hotspot cells in realm)
    perc <- setDT(as.data.table(t(round((ncell/nrow(sub_rcp))*100, 1))))
    colnames(perc) <- paste0("perc_", colnames(perc))
    perc[, ':='(Scen = rcp, Realm = realm)]
    # percentage of extreme cells rel. to realm (ncell extreme across realm/total number of cells in realm))
    percRealm <- setDT(as.data.table(t(round((ncell_realm/nrow(realm_cells))*100, 1))))
    colnames(percRealm) <- paste0("percRealm_", colnames(percRealm))
    percRealm[, ':='(Scen = rcp, Realm = realm)]
    ncell <- setDT(as.data.frame(t(ncell)))
    ncell[, ':='(Scen = rcp, Realm = realm)]
    outDT <- data.table(
        # Scenario
        "Scen" = rcp,
        # Realm
        "Realm" = realm,
        # total number of hotspot cells
        "Total_HS_cells" = as.integer(nrow(sub_rcp)),
        # total number of cells in a realm
        "Total_Realm_cells" = as.integer(nrow(realm_cells)))
    outDT <- merge(outDT, ncell, by = c("Realm", "Scen"))
    outDT <- merge(outDT, perc, by = c("Realm", "Scen"))
    outDT <- merge(outDT, percRealm, by = c("Realm", "Scen"))
    return(outDT)
  })
  return(rbindlist(realm_rcp, use.names = TRUE))
})
num_ext_cells <- rbindlist(num_ext_cells, use.names = TRUE)
num_ext_cells[, c("Realm","Scen","Total_HS_cells","Total_Realm_cells", "95%", "perc_95%","percRealm_95%")][order(Realm, Scen), ]
```

The results show that in some regions the most extreme SNR values were encountered in the past. For example the Caribbean & Gulf of Mexico bioregion under both RCP 8.5 and RCP 4.5 had more extreme conditions in the past. This is also shown by the histograms in the previous section (`04_CalculateOverlap`). However, the opposite is also true with some regions showing 100% of hotspots cells having more extreme conditions than those experienced in the past.

```{r showExtreme, echo = FALSE}
num_ext_cells[, c("Realm","Scen","Total_HS_cells","Total_Realm_cells", "95%", "perc_95%","percRealm_95%")][order(-`perc_95%`), ]
```

Here we can look at the change in hotspots cells across all scenarios

```{r hsChanges, results='markup'}
## how many HS cells are more extreme than the total number of HS cells?
## the global percentage of hotspots of marine species richness that are likely to experience no-analogue rates of 21st century climate warming
## these numbers are reported in the main text
num_ext_cells[, .(GlbCh = round(sum(`95%`)/sum(Total_HS_cells), 2)*100), by = Scen]

## weighted mean based on total realm area
num_ext_cells[, .(WtdMean = round(Hmisc::wtd.mean(x = `perc_95%`, weights = Total_Realm_cells, normwt = FALSE))),
              by = Scen]
```

## Sensitivity analysis

The results of the sensitivity analysis show that for some regions there are changes with regards to the number of hotspot cells that will experience more extreme conditions in the future depending on the threshold of past SNR

```{r sensResults, results='markup', rownames.print = TRUE, rows.print = 13}
cols <- c("Realm", "Scen", "Total_HS_cells", colnames(num_ext_cells)[grepl("^[9]", colnames(num_ext_cells))])
sens <- melt(num_ext_cells[, ..cols][order(-Scen), ], id.vars = c("Realm", "Scen"))
sens_realms <- sapply(unique(sens$Realm), function(x) {
  sumCells <- sens[Realm == x & variable == "Total_HS_cells",][["value"]][1L]
  extCells45 <- sens[Realm == x & variable != "Total_HS_cells" & Scen == "RCP45",][["value"]]
  extCells85 <- sens[Realm == x & variable != "Total_HS_cells" & Scen == "RCP85",][["value"]]
  r45 <- sd((extCells45/sumCells)*100)
  r85 <- sd((extCells85/sumCells)*100)
  return(c("RCP 4.5" = r45, "RCP 8.5" = r85))
})
# Sensitivity (sd in % cells depending on threshold)
round(sens_realms, 1)

# average sd across all realms
round(apply(as.data.frame(t(sens_realms)), 2, mean), 2)

# range in sd across all realms
round(apply(as.data.frame(t(sens_realms)), 2, range), 2)
```

